### 2018-05-18 manuscript/letter.tex
Review of the part about gravity modeling
Renames the section to just "Gravity modeling" because that is what it's
about.
Adds an introduction to the section and removes a lot of repetition to
make the text flow better.
Also change some of the language to avoid blaming people too much.

% gjilguid2e.tex
% V2.0 released 1998 December 18
% V2.1 released 2003 October 7 -- Gregor Hutton, updated the web address for the style files.

%\documentclass[extra,mreferee]{gji}
\documentclass[extra]{gji}
\usepackage{timet}

%Extra packages
\usepackage{graphicx}
\usepackage{amsmath}

\title[Gravity anomaly or gravity disturbance?]
      {Should geophysicists use gravity anomaly or gravity disturbance?}

\author[Hallam et al.]{
\parbox{\linewidth}{%
    Kristoffer A. T. Hallam$^{1}$\thanks{Emails: kristoffer.hallam@gmail.com, vanderlei@on.br},
    Vanderlei C. Oliveira Jr$^1$,
    Val\'{e}ria C. F. Barbosa$^1$, and \linebreak Leonardo Uieda$^2$
    \vspace{0.3cm}
}%
    \\
    $^1$ Department of Geophysics, Observat\'{o}rio Nacional, Rio de Janeiro, Brazil \\
    $^2$ Department of Geology and Geophysics, SOEST, University of Hawai'i at M\={a}noa, Honolulu, USA
}

\date{Received 2018 Month XX; in original form 2018 Month XX}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\volume{XXX}
\pubyear{2018}

%\def\LaTeX{L\kern-.36em\raise.3ex\hbox{{\small A}}\kern-.15em
%    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}
%\def\LATeX{L\kern-.36em\raise.3ex\hbox{{\Large A}}\kern-.15em
%    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}
% Authors with AMS fonts and mssymb.tex can comment out the following
% line to get the correct symbol for Geophysical Journal International.
\let\leqslant=\leq

\newtheorem{theorem}{Theorem}[section]

\newcommand{\versor}[1]{\hat{\mitbf{#1}}}
\renewcommand{\vector}[1]{\vec{\mitbf{#1}}}

\begin{document}

\label{firstpage}

\maketitle

\begin{summary}
 In applied geophysics, the gravity anomaly has long been used for representing the
 gravitational effect produced by gravity sources, which are defined as the density
 contrasts between the internal density distribution of the Earth and the normal Earth.
 The gravity anomaly is defined by the difference between the gravity (produced by the
 Earth) on the Geoid and the normal gravity (produced by the normal Earth)
 on reference ellipsoid, both at the same geodetic latitude and longitude.
 Because these quantities are not defined at the same point,
 the gravity anomaly does not represent the gravitational effect due to the gravity
 sources and, consequently, cannot be considered a harmonic function.
 On the contrary, the gravity disturbance is defined by the difference between
 the gravity and normal gravity at the same point. Consequently, the centrifugal
 effects can be neglected, the gravity disturbance can be considered a harmonic
 function representing the gravitational effect due to the gravity sources and then
 it satisfies the premise behind most of the processing techniques of potential-field
 data (e.g., upward/downward continuations). Then, rigorously, the gravity disturbance
 is  more appropriated to be used for estimating density variations in subsurface.
 The use the gravity anomaly for most of the geophysicists carries the
 implicit meaning that it can be considered a good approximation of the gravity
 disturbance. It is important to bear in mind, however, that in Brazil, for example,
 the maximum absolute difference between the gravity disturbance (on or close
 to the Earth’s surface) and the Free-air anomaly reaches 9.3 mGal.

\end{summary}

\begin{keywords}
 potential fields -- gravity disturbance -- gravity anomaly -- gravity modelling.
\end{keywords}

\section{Introduction}

The \textit{gravity vector} is the sum of the gravitational and centrifugal
accelerations felt by a body at rest on the Earth's surface.
Its intensity is what geoscientists call \textit{gravity}
\citep{heiskanen-moritz1967, hofmann-wellenhof-moritz2005}.
When measured on moving platforms (e.g., airplanes,
helicopters, marine vessels), there are additional
non-gravitational accelerations due to the motion of the vehicle,
such as the Coriolis acceleration and high-frequency vibrations
\citep{glennie-etal2000,nabighian-etal2005-grav,baumann-etal2012}.
Geodesists use gravity measurements to estimate the Geoid \citep{li2001}
whereas geophysicists use gravity to estimate the Earth's
internal density distribution.
Hence, geophysicists are usually only interested
in the gravitational component of the observed gravity
because that is what that reflects the Earth's internal density distribution.

The first step toward isolating the gravitational component
is to remove the effects of vehicle motion, Earth tides, instrumental drift,
and barometric pressure changes, among others.
If these effects are properly removed,
the resultant observations are considered to be solely
the sum of the centrifugal acceleration due to the Earth's rotation and
the gravitational acceleration produced by
the internal density distribution of the whole Earth.
The isolation of this particular gravitational component,
and its subsequent use for estimating density
distributions related to geological structures in subsurface,
are the main goals of applied gravimetry \citep{blakely1996}.
The computation of the gravitational effect produced by
the geological structures is known in geophysics as
\textit{gravity modelling}.
Not to be confused with \textit{gravity field modelling}, which
is commonly used in geodesy to denote the mathematical characterization of the
gravity field in local, regional, or global scales (e.g., spherical harmonic
analysis).

We present a discussion of whether the gravity disturbance or the gravity
anomaly should be used in geophysical applications.
The concepts discussed here are well-established in the literature.
However, the debate around this theoretical issue has been
carried about from a more geodetic than a geophysical point of view
\citep{lafehr1991,chapin1996,li2001,fairhead2003,
hackney-featherstone2003,hinze2005}.
We will approach the topic by comparing the geophysical interpretations of the
disturbance and the anomaly.
Our reasoning suggests that the gravity disturbance is more appropriated for
geophysical purposes than the gravity anomaly.

\section{Normal Earth and normal gravity}

The Earth's gravity field is traditionally approximated
by the gravity field of a reference ellipsoid (or level ellipsoid).
This model is rigid and geocentric,
with a minor axis $b$ which coincides with
the mean rotation axis of the Earth $Z$.
The ellipsoid has the same total mass as the Earth (including the atmosphere)
and also the same angular velocity \citep{heiskanen-moritz1967,
vanicek1987,hofmann-wellenhof-moritz2005,torge2012}.
Its limiting surface coincides with
a particular equipotential of its own gravity field.
Here, we follow \citep{torge2012} and call this reference ellipsoid
the \textit{normal Earth}.
Analogously to the Earth, the \textit{normal gravity vector} is
the sum of the
gravitational and centrifugal accelerations exerted by the normal
Earth on a body at rest at a point $P$.
The intensity of the normal gravity vector is called \textit{normal gravity}.

It is worth noting that, although the normal Earth has the same total mass as
the Earth, its internal density distribution is left unspecified.
The search for geologically meaningful density distributions
for the interior of the normal Earth has
geophysical rather than geodetic motives \citep{marussi1974}.
In physical geodesy, a model for the normal gravity field
can be arbitrarily defined with the sole purpose of
keeping the difference from the actual gravity field
as small as possible \citep{vanicek1987}.
The only condition imposed on its internal density
distribution is that it produces a gravity field
having a particular equipotential which coincides
with its limiting surface.

\section{Terrestrial Reference Systems used in gravity modelling}

For geophysical gravity modelling purposes, there are three important
Terrestrial Reference Systems.
They rotate with the Earth and are used for describing
positions and movements of objects on and close to the Earth's surface
\citep{torge2012}.

The first is a geocentric system of Cartesian coordinates
having the $Z$-axis coincident with the mean rotational axis of the Earth,
the $X$-axis pointing to the Greenwich meridian,
and the $Y$-axis completing a right-handed system (Fig. \ref{fig:GCS-GGS}).
This reference system is called by different names in the literature:
Mean Terrestrial System \citep[e.g.,][]{soler1976},
Earth-fixed geocentric Cartesian system \citep[e.g.,][]{torge2012}
or Earth-centered Earth-fixed system \citep[e.g.,][]{bouman_etal2013},
for example.
Here, we opted for using the term Geocentric Cartesian System (GCS).

The second is a geocentric system of geodetic coordinates
(the Geocentric Geodetic System or GGS),
which is defined by the reference ellipsoid used as the Normal Earth
\citep{heiskanen-moritz1967, soler1976, torge2012, bouman_etal2013}.
The position of a point is defined by
the \textit{geometric height} $h$,
the \textit{geodetic latitude} $\varphi$,
and the \textit{longitude} $\lambda$ (Fig. \ref{fig:GCS-GGS}).
There are also three mutually-orthogonal unit vectors defined at a given point
\citep{soler1976}:

\begin{equation}
    \versor{u} =
    \begin{bmatrix}
        \cos\varphi \, \cos\lambda \\
        \cos\varphi \, \sin\lambda \\
        \sin\varphi
    \end{bmatrix}
    \versor{v} =
        \begin{bmatrix}
        -\sin\varphi \, \cos\lambda \\
        -\sin\varphi \, \sin\lambda \\
        \cos\varphi
    \end{bmatrix}
    \versor{w} =
    \begin{bmatrix}
        -\sin\lambda \\
        \cos\lambda \\
        0
    \end{bmatrix} .
    \label{eq:unit-vectors}
\end{equation}

\noindent Equations for converting between these systems can be found in
the literature \citep[e.g.,][]{heiskanen-moritz1967, torge2012,
bouman_etal2013}.

The third is the Topocentric Cartesian System (TCS),
which is commonly used in local or regional scale geophysical studies.
The origin of the TCS is located at a point $P$,
axes $x$ and $y$ are parallel to
the unit vectors $\versor{v}$ and $\versor{w}$
(eq. \ref{eq:unit-vectors}), respectively,
and the $z$-axis is opposite to the unit vector $\versor{u}$
(Fig. \ref{fig:TCS}).

\section{gravity disturbance and gravity anomaly}

Let $\vector{\gamma}_{P}$ and $\vector{g}_{P}$ and be, respectively,
the \textit{normal gravity vector} and the Earth's \textit{gravity vector}
(corrected from non-gravitational effects due to vehicle
motion and time variations such as Earth tides and instrumental drift),
both located at a point $P$.
In this case, the gravity vector $\vector{g}_{P}$ is the
gradient of the scalar \textit{gravity potential},
which is the sum of a gravitational potential and a centrifugal potential.
Similarly, the normal gravity vector $\vector{\gamma}_{P}$ is the
gradient of the scalar \textit{normal potential},
which is also the sum of a gravitational and a centrifugal potential.
By definition, the centrifugal part of
the normal potential and the gravity potential are equal.

The difference between $\vector{g}_{P}$ and $\vector{\gamma}_{P}$,
which we emphasize are located at the same point $P$,
is the \textit{gravity disturbance vector} (Fig. \ref{fig:surfaces}):

\begin{equation}
\delta \vector{g}_{P} =
\vector{g}_{P} - \vector{\gamma}_{P} \: .
\label{eq:gravity-disturbance-vector}
\end{equation}

Because the centrifugal parts of both the normal gravity vector
and the Earth's gravity vector are equal,
the gravity disturbance vector $\delta \vector{g}_{P}$
represents a purely gravitational effect.
As a consequence of the superposition principal \citep{blakely1996},
this effect is caused by contrasts between
the internal density distributions
of the actual Earth and the normal Earth.
In applied geophysics, these density differences are generally called
\textit{anomalous masses} \citep[e.g.,][]{hammer1945,lafehr1965},
\textit{density anomalies} \citep[e.g.,][]{forsberg1984},
or \textit{gravity sources} \citep[e.g.,][]{blakely1996}.
Here, we opted for using the last term.

The difference between the magnitudes of the gravity vector
$g_{P} = \| \vector{g}_{P} \|$ and the normal gravity vector
$\gamma_{P} = \| \vector{\gamma}_{P} \|$,
at the same point $P$, is the \textit{gravity disturbance}
\citep{heiskanen-moritz1967, hofmann-wellenhof-moritz2005}:

\begin{equation}
\delta g_{P} = g_{P} - \gamma_{P} \: .
\label{eq:gravity-disturbance}
\end{equation}

\noindent
Notice that $\delta g_{P}$ is not equivalent
to the magnitude of the gravity disturbance vector
$\delta \vector{g}_{P}$ \citep{barthelmes2013, sanso_sideris2013}.

The \textit{gravity anomaly vector}
is the difference between the gravity
vector at a point $Q^{\prime}$ on the geoid
(a particular equipotential surface of the gravity potential)
and the normal gravity vector at a point $Q$ on the surface of the ellipsoid,
both located at the same geodetic latitude and longitude
(Fig. \ref{fig:surfaces}):

\begin{equation}
\Delta \vector{g} = \vector{g}_{Q^{\prime}} - \vector{\gamma}_{Q} .
\label{eq:gravity-anomaly-vector}
\end{equation}

Similarly to the gravity disturbance (eq. \ref{eq:gravity-disturbance}),
the \textit{gravity anomaly} is given by

\begin{equation}
\Delta g = g_{Q^{\prime}} - \gamma_{Q} ,
\label{eq:gravity-anomaly}
\end{equation}

\noindent
where $g_{Q^{\prime}}$ is the magnitude of the gravity vector on the geoid,
and $\gamma_{Q}$ is the magnitude of the normal gravity vector
on the reference ellipsoid (Fig. \ref{fig:surfaces}).

The gravity anomaly is not defined at a particular point.
Instead, as properly pointed out by \citet{barthelmes2013},
it depends only on longitude and latitude and is not a function of height.
Because gravity and normal gravity are located at different points,
the gravity anomaly is a combination of gravitational and centrifugal effects,
rather than purely gravitational effects like the gravity disturbance.
Consequently, its upward continuation, for example, cannot be rigorously done
because it is not a harmonic function.
However, upward continuation of gravity anomalies is commonly seen in the
literature.
In doing so, an implicit assumption is being made that the gravity anomaly is
an approximation of the gravity disturbance,
which in turn can be represented by a harmonic function and can be upward
continued.

\section{The geophysical meaning of gravity reductions}

Different approximations for the gravity anomaly can be calculated
depending on the corrections applied to them.
These corrections are usually called \textit{gravity reductions}
and their aim is to approximate the magnitude of gravity at the geoid
$g_{Q^\prime}$ from the magnitude of gravity $g_P$ measured at a point $P$
on or above the surface of the Earth.

The \textit{free-air anomaly}, for example, can be defined as
\citep{blakely1996, hofmann-wellenhof-moritz2005}:

\begin{equation}
\Delta g^{F}
= \left(g_{P} - \frac{\partial \gamma}{\partial h} \, H_{P} \right)
- \gamma_{Q} \: ,
\label{eq:free-air-anomaly}
\end{equation}

\noindent
where $\partial \gamma/\partial h \approx -0.3086$ mGal/m is the
derivative of the normal gravity with respect to the geometric height $h$
and $H_{P}$ is the orthometric height (i.e., with respect to the geoid)
of point $P$ (Fig. \ref{fig:surfaces}).
The term between parenthesis represents an
approximation of the gravity at the point $Q^{\prime}$ on the geoid,
disregarding the effect of topographic masses.
Similarly, we can apply the same correction
to calculate the gravity disturbance by
approximating the normal gravity $\gamma_{P}$
at a point $P$ on the Earth's surface
from the normal gravity at the ellipsoid $\gamma_{Q}$:

\begin{equation}
\delta g_{P} \approx g_{P} -
\left( \gamma_{Q} + \frac{\partial \gamma}{\partial h} h_P \right) \: .
\label{eq:gravity-disturbance-approx}
\end{equation}

\noindent
where $h_P$ is the geometric height of point P.
Such an approximation is usually not required because
an analytical expression exists for calculating $\gamma$ at any point
on or above the ellipsoid \citep{li2001}.
The same cannot be said for the free-air anomaly because
there is usually no way of measuring gravity at the geoid ($g_{Q^\prime}$).

Using eqs. \ref{eq:free-air-anomaly} and \ref{eq:gravity-disturbance-approx},
we can estimate the absolute difference between the gravity disturbance and the
free-air anomaly:

\begin{equation}
\left\vert \delta g_{P} - \Delta g^{F} \right\vert \approx
\left\vert \frac{\partial \gamma}{\partial h} N \right\vert \: ,
\label{eq:difference-disturbance-free-air}
\end{equation}

\noindent
where $N \approx h - H$ is the geoidal height (Fig. \ref{fig:surfaces}).
This approximation assumes that the geoid and the surface of
the reference ellipsoid are close to being parallel at $P$
and the surrounding area.
Eqs \ref{eq:gravity-disturbance-approx}
and \ref{eq:difference-disturbance-free-air} are commonly used
in geodesy to define the \textit{fundamental equation of physical
geodesy} \citep{hofmann-wellenhof-moritz2005}.

We know empirically that $N$ is $\approx \pm 1$ m on the oceans
and reaches a maximum absolute value of $\approx 120$ m
\citep[e.g.,][]{torge2012, sanso_sideris2013}.
In Brazil, for example, the geoidal height
varies between $\approx \pm 30$ m \citep{ibge_mapgeo2015}.
Consequently,
the maximum absolute difference between the gravity disturbance and
the free-air anomaly in Brazil,
according to eq. \ref{eq:difference-disturbance-free-air},
is $\approx 9.258$ mGal.

The \textit{Bouguer anomaly}
has the aim of removing the gravitational effect of the topographic masses
between the observation point $P$ and the point on the geoid $Q^\prime$.
It is defined in the continents as
\citep{blakely1996, hofmann-wellenhof-moritz2005}:

\begin{equation}
\Delta g^{B} =
\left(g_{P}
    - \frac{\partial \gamma}{\partial h} \, H_{P}
    - 2 \pi G \rho_{t} \, H_{P} \right)
- \gamma_{Q} \: ,
\label{eq:simple-bouguer-anomaly}
\end{equation}

\noindent
where $\rho_t$ is the density, $G$ is the gravitational constant,
and the rightmost term in parenthesis is the simple Bouguer correction.
This term approximates the gravitational effect that all topographic
masses above the geoid exert on point $P$ by the effect of a homogeneous,
infinitely extended slab of constant density $\rho_{t}$ and thickness $H_{P}$.
The same correction can be applied to the gravity disturbance
(using the approximation in eq. \ref{eq:gravity-disturbance-approx})
with the aim of removing the effect of topographic masses between
the observation point $P$ and a point on the ellipsoid $Q$.
We will call this quantity the \textit{Bouguer disturbance}:

\begin{equation}
\delta g_P^{B} \approx
g_{P} -
\left( \gamma_{Q} + \frac{\partial \gamma}{\partial h} h_P \right)
- 2 \pi G \rho_{t} \, h_{P} \: .
\label{eq:bouguer-disturbance}
\end{equation}

At first glance, the only difference between eqs.
\ref{eq:simple-bouguer-anomaly} and \ref{eq:bouguer-disturbance} is the use of
orthometric ($H_P$) versus geometric ($h_P$) heights.
However, their objectives and geophysical interpretations are significantly
different.
The corrections applied in eq. \ref{eq:simple-bouguer-anomaly} approximate
gravity at the geoid from measurements made on the Earth's surface by removing
the effects of the distance and the topographic masses between the two points.
Apart from the centrifugal effect mentioned in the previous section,
the gravitational effect of topographic masses between the geoid and ellipsoid
are still present in the anomaly.
Conversely, when calculating the Bouguer disturbance one removes from
the observed gravity the effects of an ellipsoidal reference model (the normal
Earth) and the known topographic masses between the surface of the ellipsoid
and observation point.
Hence, the Bouguer disturbance can be interpreted as being solely the
gravitational effect produced by the density contrasts between the actual
internal density distribution of the Earth and that of the normal Earth.

There was a certain lack of comprehension regarding the geophysical meaning of
gravity anomalies until the mid 90s.
As properly pointed out by \citet{chapin1996},
``although the corrections which bring about a Bouguer gravity anomaly are well
established, the reasons for doing them are not well understood. One cause of
this common misunderstanding is that the subject has been poorly presented in
many of the basic texts''.
In his seminal book, \citet{blakely1996} shed some light on the geophysical
meaning of gravity anomalies from the perspective of applied geophysics and
their relationship to the gravity sources.
We complement this interpretation by stressing the fact that
gravity anomalies reflect not only the gravity sources,
but also a combination of gravitational and centrifugal effects.
These additional, non-harmonic and undesired effects are due to the calculation
of the normal gravity at a point other than that were gravity is measured.
In practice, this issue can be avoided entirely by performing the free-air and
Bouguer corrections using geometric instead of orthometric heights.

\section{Gravity modelling}

We have established that the gravity disturbance is due solely to the
gravitational effect of the gravity sources.
Now we will discuss how this relates to some assumptions made during gravity
modelling.

The gravity vector $\vector{g}_i$ (eq. \ref{eq:gravity-disturbance-vector})
at a point $(x_i, y_i, z_i)$ in the TCS (Fig. \ref{fig:TCS}a) can be
represented by:

\begin{equation}
\vector{g}_i = \vector{\gamma}_i + \delta \vector{g}_i \: .
\label{eq:gravity-vector-TCS}
\end{equation}

\noindent
Furthermore, the gravity $g_i$ can be approximated by a first order Taylor's
expansion \citep{sanso_sideris2013}:

\begin{equation}
g_i \approx \gamma_i +
\versor{\gamma}_i^{\top} \delta \vector{g}_i \: ,
\label{eq:gobs-approx}
\end{equation}

\noindent
where $^{\top}$ denotes transposition and $\versor{\gamma}_{i}$ is a unit
vector with the direction of the normal gravity vector.
This approximation is valid because
$\gamma_{i} \gg \| \delta \mathbf{g}_{i} \|$ at all points on or above the
Earth's surface.
It is known in geodesy \citep[e.g.,][]{sanso_sideris2013}
and is analogous to the approximation used for
total-field magnetic anomalies \citep[e.g.,][]{blakely1996}.
In local or regional scales, the unit
vector $\versor{\gamma}_{i}$
may be considered constant throughout the study area and
parallel to the $z$ axis of the TCS (Fig. \ref{fig:TCS}a).

Using eq. \ref{eq:gobs-approx} and fixing $\gamma$ at a point $P$ at
the origin of the TCS (Fig. \ref{fig:TCS}a),
the gravity disturbance can be rewritten as:

\begin{equation}
\delta g_{i} \approx \versor{\gamma}_{P}^{\top} \delta \vector{g}_{i} \: .
\label{eq:gravity-disturbance-approx-TCS}
\end{equation}

\noindent
This equation shows that the gravity disturbance $\delta g_{i}$ is
approximately the component of the $\delta \vector{g}_{i}$ in the direction of
the normal gravity vector $\vector{\gamma}_P$
\citep{hofmann-wellenhof-moritz2005, sanso_sideris2013}.
Because $z$-axis of the TCS is parallel to $\vector{\gamma}_P$
(Fig. \ref{fig:TCS}b), the gravity disturbance can be defined as the
$z$-component of the gravitational attraction exerted by the gravity sources at
the point $(x_{i}, y_{i}, z_{i})$.
As a consequence, the gravity disturbance can be modelled by the harmonic
function:

\begin{equation}
d_{i} = \iiint\limits_{v}
\frac{G \, \Delta\rho(x^{\prime}, y^{\prime}, z^{\prime}) \,
[z_{i} - z^{\prime}]}
{\sqrt{[x_{i} - x^{\prime}]^{2} +
[y_{i} - y^{\prime}]^{2} + [z_{i} - z^{\prime}]^{2}}} \: dv^{\prime} \: ,
\label{eq:gz-local}
\end{equation}

\noindent
where $\Delta\rho(x^{\prime}, y^{\prime}, z^{\prime})$
is the density contrast at a point $(x^{\prime}, y^{\prime}, z^{\prime})$
within the volume $v$ and the integration is conducted over $x^{\prime}$,
$y^{\prime}$ and $z^{\prime}$.

In practice, most instances of gravity modelling in the literature use Bouguer
anomalies and function $d_{i}$ (usually presented as ``$g_z$'') to represent
the gravitational effect of gravity sources.
Consequently, these studies are using the gravity
anomaly as an approximation for the gravity disturbance.
Notice that $d_{i}$ does not depend on the
height with respect to the geoid (orthometric height).
Rather, it depends on the relative position of the gravity sources and the
observation point $(x_{i}, y_{i}, z_{i})$ in the TCS (Fig. \ref{fig:TCS}a).
Finally, it is important to mention that gravimeters do not measure $d_{i}$ but
$\| \vector{g}_i \|$.

\section{Conclusions}

Although the gravity disturbance is a well-known quantity in geodesy,
its use in the applied geophysics is uncommon.
For years, geophysicists have been using gravity anomaly for modelling
the gravitational effects due to geological bodies.
The question "Should geophysicists use gravity anomaly or gravity
disturbance for modelling the gravitational effect of gravity sources?"
is of more than academic interest.

Theoretically, the gravity anomaly is the difference between the
gravity on the Geoid and the normal gravity on the reference ellipsoid,
both at the same geodetic latitude and longitude.
On the other hand, the gravity disturbance is defined by the difference
between gravity and normal gravity at the same point.
Because these quantities are at the same point, the gravity disturbance
is a purely gravitational effect due to the gravity sources whereas the
gravity anomaly is a combination of gravitational and centrifugal
effects.
Consequently, the gravity disturbance is a harmonic function whereas the
gravity anomaly is not.
Because the gravity disturbance is a purely gravitational effect, it can
be represented by a harmonic function and satisfies the premise required
by the most techniques for processing potential-field data (e.g.,
upward/downward continuation, data processing with equivalent layer,
conversions between gravity and magnetic data, computation of vertical
derivatives via Fourier and Hilbert transforms).
Hence, the gravity disturbance is the more appropriated quantity for
representing the gravity effect produced by the gravity sources.

In practical problems of applied geophysics, geophysicists perform the
gravity-forward modelling by computing the vertical component of the
gravitational attraction exerted by the sources at the observation
points, regardless of the Geoid or reference ellipsoid positions,
for example.
This means that, actually, they compute gravity disturbance to fit
a gravity anomaly (e.g., Bouguer anomaly) and this is a clear inconsistency.
Moreover, upward or downward continuations, for example, are commonly applied to
gravity anomalies. In this case, the gravity anomaly is wrongly considered a
harmonic function and this is another inconsistency.

By adopting these practical procedures, geophysicists are implicitly
taking the gravity anomaly as a good approximation of the gravity
disturbance. But why should geophysicists care about this usual
procedure? Because it may result in wrong interpretations of
quantitative nature. In Brazil, for example, the maximum difference
between the gravity disturbance and the free-air anomaly reaches 9.3
mGal. Such difference on the Earth's surface may lead, for example, to
an error of about xxx m in the depth-to-center estimate of a sphere-like
source with true depth-to-center at xxx m, radius xxxx m and density
contrast of xxx $kg \, m^{-3}$.

\begin{acknowledgments}
The authors would like to thank the editor and all the
reviewers for their criticisms and corrections.
\end{acknowledgments}

\bibliographystyle{gji}
\bibliography{bib-file}

\appendix
%\section{For authors}
%
%Table~\ref{authors} is a list of design macros which are unique to GJI. The
%list displays each macro's name and description.
%
%\section{For editors}
%
%The additional features shown in Table~\ref{editors} may be used for
%production purposes.
%
%\bsp % ``This paper has been produced using the Blackwell
%     %   Publishing GJI \LaTeXe\ class file.''

\begin{figure}
%    \vspace{5.5cm}
    \includegraphics{figures/cartesian-geodetic-systems.png}
    \caption{Schematic representation of the Geocentric Cartesian System (GCS) and the Geocentric Geodetic System (GGS). The GCS has the $Z$-axis coincident with the mean Earth's rotational axis,
    the $X$-axis pointing to the Greenwich meridian and the $Y$-axis
    directed so as to obtain a right-handed system. The GGS is      defined by an oblate ellipsoid with semi-minor axis $b$,
    coincident with the $Z$-axis of GCS, and a semi-major
    axis $a$. In this coordinate system, the position of a point is
    determined by the geometric height $h$, geodetic latitude
    $\varphi$ and longitude $\lambda$.
    The Earth's center of mass is represented
    by $O$, $P$ represents a point $(h_{P}, \varphi_{P}, \lambda_{P})$ and $P^{\prime}$ its projection onto the plane $XY$
    (Equatorial plane). The plane containing $O$, $P$ and
    $P^{\prime}$ is represented in gray in (a) and (b).
    The unit vectors $\versor{u}_{P}$, $\versor{v}_{P}$ and
    $\versor{w}_{P}$ define three mutually orthogonal
    directions at $P$ (eq. \ref{eq:unit-vectors}).
    In (b), $Q$ represents a point $(h_{Q}, \varphi_{Q}, \lambda_{Q})$,
    which is the projection of $P$ onto the reference ellipsoid, at
    the same latitude and longitude ($\varphi_{Q} = \varphi_{P}$ and $\lambda_{Q} = \lambda_{P}$).}
  \label{fig:GCS-GGS}
\end{figure}

\begin{figure}
    \includegraphics{figures/local-system.png}
    \caption{(a) Schematic representation of a topocentric Cartesian
    coordinate system (TCS) with origin at a point $P$. The axes $x$
    and $y$ are parallel to the unit vectors $\versor{v}_{P}$
    and $\versor{w}_{P}$ (eq. \ref{eq:unit-vectors} and
    Fig. \ref{fig:GCS-GGS}), respectively. On the other hand, the $z$ axis is opposite to the unit vector $\versor{u}_{P}$ (eq.
    \ref{eq:unit-vectors} and Fig. \ref{fig:GCS-GGS}) and points downward. The gray plane is the same shown in Fig. \ref{fig:GCS-GGS}.
    (b) Schematic representation of the gravity vector
    $\vector{g}_{P}$, normal gravity vector $\vector{\gamma}_{P}$,
    gravity disturbance vector $\delta \vector{g}_{P}$
    (eq. \ref{eq:gravity-disturbance-vector}) and unit vector
    $\versor{u}_{P}$ (eq. \ref{eq:unit-vectors}) at a point
    $P$.}
  \label{fig:TCS}
\end{figure}

\begin{figure}
    \includegraphics{figures/surfaces.png}
    \caption{Schematic representation of the gravity vector
        $\vector{g}_{P}$, normal gravity vector $\vector{\gamma}_{P}$,
        gravity disturbance vector $\delta \vector{g}_{P}$
        (eq. \ref{eq:gravity-disturbance-vector}), unit vector
        $\versor{u}_{P}$ (eq. \ref{eq:unit-vectors}) at a point $P$
        on the surface of the Earth, gravity vector $\vector{g}_{Q^{\prime}}$
        at a point $Q^{\prime}$ on the Geoid, normal gravity vector
        $\vector{\gamma}_{Q}$ at a point $Q$ on the reference ellipsoid,
        geometric height $h$, orthometric height $H$ and geoidal undulation
        $N$ \citep{heiskanen-moritz1967}.
        The dashed line passing through
        $Q$, $Q^{\prime}$ and $P$ is normal to the
        surface of the reference ellipsoid at $Q$. This figure
        shows a commonly used approximation in which the ellipsoid
        surface and the Geoid are represented as parallel surfaces,
        so that $h \approx H + N$.}
  \label{fig:surfaces}
\end{figure}

\label{lastpage}

\end{document}